Get EPA ecoregions data

Tyler L. McIntosh
CU Boulder CIRES Earth Lab
Last updated: 1/24/24

This script uses the following naming conventions wherever possible:
 lowerCamelCase for variables
 period.separated for functions
 underscore_separated for files

# Setup workspace
User-set global parameters
Directory management
Package management

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

rm(list=ls()) #Ensure empty workspace if running from beginning

#################################################
#####EVERYTHING HERE SHOULD BE SET MANUALLY######
#################################################

computing <- "local" #Sets computing location and file setup; "cyverse" or "local"

#################################################

# Directory management ----
require(here)

#Set here location
here::i_am("code/00_download_epa_regions.qmd")

#Set directories
if(computing == "local") {
  home <- here::here()
} else if(computing == "cyverse") {
  home <- "~/data-store/data/iplant/home/shared/earthlab/macrosystems/disturbance-stack-patch-analysis"
}

#Raw data
rawDir <- file.path(home, 'data/raw')
if (!dir.exists(rawDir)){
  dir.create(rawDir)
}

# Package management ----

source(here::here('code/functions.R'), local = FALSE)

#Check the required libraries and download if needed
packageList <- c()

install.and.load.packages(packageList, autoInstall = "y")

```

# Download EPA ecoregions data directly from URL

```{r download-epa}
#Avoid download timeout
options(timeout = max(1000, getOption("timeout")))

#URLs for downloads
epaUrls <- c("https://gaftp.epa.gov/EPADataCommons/ORD/Ecoregions/us/us_eco_l3.zip",
             "https://gaftp.epa.gov/EPADataCommons/ORD/Ecoregions/us/us_eco_l4.zip")
destFiles <- file.path(rawDir, basename(epaUrls))

#Download
mapply(FUN = function(url, destfile) {download.file(url = url,
                                          destfile = destfile,
                                          mode = "wb")},
       url = epaUrls,
       destfile = destFiles)

#Unzip downloaded files
mapply(FUN = function(destfile, exdir) {unzip(zipfile = destfile,
                                              files = NULL,
                                              exdir = exdir)},
       destfile = destFiles,
       exdir = gsub(pattern = ".zip", replacement = "", x = destFiles))

```

#Acquire EPA ecoregions via VSIZIP/VSICURL (NOT USED IN THIS WORKFLOW)

```{r}

# require(glue)
# require(sf)
# 
# test <- glue::glue(
#   "/vsizip/vsicurl/", #magic remote connection
#   "https://gaftp.epa.gov/EPADataCommons/ORD/Ecoregions/us/us_eco_l3.zip", #copied link to download location
#   "/us_eco_l3.shp") |> #path inside zip file
#   sf::st_read()
# test2 <- glue::glue(
#   "/vsizip/vsicurl/", #magic remote connection
#   "https://gaftp.epa.gov/EPADataCommons/ORD/Ecoregions/us/us_eco_l4.zip", #copied link to download location
#   "/us_eco_l4_no_st.shp") |> #path inside zip file
#   sf::st_read()



```

