Analyze and visualize outputs from scripts 00 & 01

Tyler L. McIntosh
CU Boulder CIRES Earth Lab
Last updated: 12/21/23

This script uses the following naming conventions wherever possible:
 lowerCamelCase for variables
 period.separated for functions
 underscore_separated for files


# Set libraries

```{r}

#Select library chunks from R_Libraries.txt and add to "list.of.packages"

######Utility libraries
utility <- c(
 	"tictoc", #benchmarking
	"beepr", # beep() will make a noise. options(error = beep) will beep on any errors
	"zip", #For zipping files (does not require external tools). mode = "cherry-pick" to select specific files w/o mirroring directory structure
	"units", #for managing units
	"remotes", #to access github
	"here", #Relative path best practices
	"profvis" #Profiling of code for efficiency and identifying bottlenecks
)


#####Data management libraries
dataManagement <- c(
	"tidyverse", #Includes: dplyr, forcats, ggplot2, lubridate, purrr, readr, stringr, tibble, tidyr
	"glue", #Easy string management
	"data.table" #More efficient dataframes and csv read/write
)

######GitHub libraries - any library from GitHub that needs to be installed via 'remotes'
github <- c(
	"tomroh/ecoregions" #pull EPA ecoregion polygons
)


#####Visualization libraries
viz <- c(
	"ggplot2",
	#"ggmisc", #various helpful tools, e.g. adding R^2 to plots
	"scales", #easily add commas to visualizations
	"gridExtra", #Multiframe plots option
	"patchwork", #Extends use of ggplot2 for multi-plot layout
	"kableExtra", #Table creation
	"RColorBrewer", #color palettes.
  # + scale_fill_brewer() for box plot, bar plot, violin plot, dot plot, etc
  # + scale_color_brewer() for lines and points
	"ggthemes", #Themes for ggplot2 package
	"lattice", #Easy visualization of multidimensional data, e.g. scatterplot matrix: splom()
	"scatterplot3d" #Visualization of three-dimensional point cloud. Example below:
	#prepare 3-D scatterplot
	#s3d <-with(trees, scatterplot3d(Girth, Height, Volume, pch=16, highlight.3d = TRUE, angle=60))
	#fit <-lm(Volume ~ Girth + Height, data  =trees) # MLR regression model
	#s3d$plane3d(fit) #produce 3-D plot/ of "best fit" plane through the 3D scatterplot
)

#####Geographic visualization
geographicViz <- c(
	"mapview", #quick interactive mapping
	"tmap", #Thematic mapping
	"tmaptools", #Supports tmap
	"leaflet" #For interactive web mapping
)


#ADD LIBRARY CHUNKS HERE
list.of.packages <- c(utility, dataManagement, github, viz, geographicViz)

```



# Setup libraries & workspace

```{r}

#Install all CRAN packages that aren't installed yet
install.packages(setdiff(list.of.packages, rownames(installed.packages())))

#Install all packages from GitHub (if any) using remotes package that aren't installed yet
if(exists("github")) { #If there are github-dependant packages...
  library(remotes) #load remotes package
  for(package in github) {
    if(! package %in% rownames(installed.packages())) {
      remotes::install_github(package) #Install package
    }
  }
}

#Replace GitHub package names in list.of.packages with the actual name of the package to load
github_names <- sub(".*/", "", github)  # Extract names after "/"
list.of.packages <- gsub(paste(github, collapse = "|"), github_names, list.of.packages)

#Load all packages
invisible(lapply(list.of.packages, library, character.only = TRUE)) #use 'invisible' to prevent output being displayed


# Clean workspace & set up environment
here::here() #Check here location
here::i_am("code/01_SpatStats.Rmd")
#OR
#setwd() #Set working directory directly
options(digits = 6) #Set standard decimal print output
options(scipen = 999) #Turn scientific notation on and off (0 = on, 999 = off)

```

# Create necessary directories - delete any unnecessary

```{r}
# Figures directory
figsDir <- here::here("figs")
if (!dir.exists(figsDir)){
  dir.create(figsDir)
}

#Derived data directory
datDerDir <- here::here("data/derived")
if (!dir.exists(datDerDir)){
  dir.create(datDerDir)
}

```


# Read in data

```{r}


#A function to get all of the sets that have been generated
fetch.sets <- function(dir) {
  fileList <- list.files(dir, pattern = ".csv", full.names = FALSE)
  setNms <- lapply(fileList, function(fl){stringr::str_split(fl, "_") |> dplyr::first() |> dplyr::first()})
  return(unique(setNms))
}

#A function to read in all csvs output with the same set name
read.set.csvs.as.one <- function(setNm, dir) {
  fileList <- list.files(dir, pattern = glue::glue("^{setNm}_"), full.names = TRUE)
  csvs <- fileList |> purrr::map(read_csv)
  oneTbl <- dplyr::bind_rows(csvs)
  return(oneTbl)
}

#A function to read all sets stored in a directory into combined dataframes, using the prior two functions
#Will return a list of all sets in the directory
read.in.all.sets <- function(dir) {
  setNms <- fetch.sets(dir)
  sets <- lapply(setNms, read.set.csvs.as.one, dir = here::here('data', 'derived', 'metric_tables'))
  names(sets) <- setNms
  return(sets)
}

#READ IN SPATSTATS OUTPUT
allDat <- read.in.all.sets(here::here('data', 'derived', 'metric_tables'))


#Load EPA ecoregion data from ecoregions package
epaL3 <- ecoregions::ContinentalUsEcoregion3
epaL4 <- ecoregions::ContinentalUsEcoregion4

  
#EPA ecoregion distinct name table for joins (from script 00_ForestStats.rmd)
nameJoin <- read_csv(here::here('data', 'derived', "epal3l4_name_join.csv"))



```

# Split into data from single and multi-layer rasters; visualize differently

```{r}

# A function to test the number of layers in each dataset and return the type specified
#dats should be a dataset read in by read.in.all.sets 
#type should be a character string of either "single" or "multi"
test.dat.layers <- function(dats, type) {
  singleLayerOrNot <- lapply(dats, FUN = function(x) {length(unique(x$layerName)) == 1})
  singleLayerInd <- which(singleLayerOrNot == TRUE)
  multiLayerInd <- which(singleLayerOrNot == FALSE)
  if(type == "single") {
    return(dats[singleLayerInd])
  } else if(type == "multi") {
    return(dats[multiLayerInd])
  } else {
    stop("'type' parameter must be either 'multi' or 'single' as a string")
  }
}

#Use test.dat.layers to split into single and multi-layer data
singleLayerDats <- test.dat.layers(allDat, type = "single")
multiLayerDats <- test.dat.layers(allDat, type = "multi")

```


# Visualize!

```{r}


```

# FROM HERE DOWN IS OLD CODE PULLED FROM ANOTHER SCRIPT - TO ADAPT


plot the metrics:
-merge the outputs into one dataframe for each metric set
-bar plot of 

AREA

NUMBER OF PATCHES

PATCH AREA MEAN & MEDIAN

PATCH CORE AREA

ED: Area-weighted edge density. The metric describes the configuration of the
landscape, e.g. because an aggregation of the same class will result in a low edge density. The metric is standardized to the total landscape area, and therefore comparisons among landscapes with different total areas are possible. Increases from 0, ED=0 if only one patch

SHAPE: Shape metric. Equals SHAPE_MN = 0 if all patches are squares. Increases, without limit, as the shapes of patches become more complex. (better than perimiter-area ratios)

CONTIG: Shape metric. Equals the mean of the contiguity index on class level for all patches.  Asses the spatial connectedness (contiguity) of cells in patches. Ranges from 0 to 1

COHESION: COHESION is an ’Aggregation metric’. It characterises the connectedness of patches belonging to
class i. It can be used to asses if patches of the same class are located aggregated or rather isolated
and thereby COHESION gives information about the configuration of the landscape. Approaches COHESION = 0 if patches of class i become more isolated. Increases
if patches of class i become more aggregated.



```{r}

tblDir <- here::here('data', 'intermediate', 'l4_tbl')

cTbls <- list.files(here::here(tblDir, 'class_metrics'), pattern = ".csv", full.names = TRUE)
pTbls <- list.files(here::here(tblDir, 'patch_metrics'), pattern = ".csv", full.names = TRUE)

classStats <- lapply(cTbls, function(f){read.csv(f)}) %>% 
  bind_rows() %>%
  filter(class == 1)
woNmC <- classStats %>%
  filter(is.na(US_L3NAME)) %>%
  select(-US_L3NAME) %>%
  left_join(nameJoin)
classStats <- classStats %>%
  filter(!is.na(US_L3NAME)) %>%
  rbind(woNmC) %>%
  filter(US_L3NAME != "Northern Basin and Range") %>%
  mutate(US_L3NAME = factor(US_L3NAME, level = c('Middle Rockies',
                                                 'Southern Rockies',
                                                 'Idaho Batholith',
                                                 'Sierra Nevada')))

patchStats <- lapply(pTbls, function(f){read.csv(f)}) %>% 
  bind_rows() %>%
  filter(class == 1)
woNmP <- patchStats %>%
  filter(is.na(US_L3NAME)) %>%
  select(-US_L3NAME) %>%
  left_join(nameJoin)
patchStats <- patchStats %>%
  filter(!is.na(US_L3NAME)) %>%
  rbind(woNmP) %>%
  filter(US_L3NAME != "Northern Basin and Range") %>%
  mutate(US_L3NAME = factor(US_L3NAME, level = c('Middle Rockies',
                                                 'Southern Rockies',
                                                 'Idaho Batholith',
                                                 'Sierra Nevada')))


```

```{r}
#PATCH STATS

ggplot(patchStats, aes(x = flNm, y = value, fill = US_L3NAME)) +
  geom_boxplot() +
  facet_wrap(~name, scales = "free") +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="black", fill="black")

ggplot(patchStats %>% filter(name == "patch area" | name == "core area"), aes(x = flNm, y = value)) +
  geom_boxplot(aes(color = US_L3NAME)) +
  facet_wrap(~name) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="black", fill="black") +
  coord_cartesian(xlim = NULL, ylim = (c(0, 250))) +
  labs(title = "Patch Areas")

ggplot(patchStats %>% filter(name == "core area index"), aes(x = flNm, y = value)) +
  geom_boxplot(aes(color = US_L3NAME)) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="black", fill="black") +
  labs(title = "Core area index")

ggplot(patchStats %>% filter(name == "shape index"), aes(x = flNm, y = value)) +
  geom_boxplot(aes(color = US_L3NAME)) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="black", fill="black") +
  labs(title = "Shape index")

#CLASS STATS

ggplot(classStats) +
  geom_col(aes(x = flNm, y = value, fill = US_L3NAME)) +
  facet_wrap(~name, scales = "free")

ggplot(classStats %>% filter(name == "largest patch index")) +
  geom_col(aes(x = flNm, y = value, fill = US_L3NAME)) +
  facet_wrap(~name, scales = "free")

ggplot(classStats %>% filter(name == "largest patch index"), aes(x = US_L3NAME, y = value)) +
  geom_boxplot(aes(color = US_L3NAME)) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="black", fill="black") +
  coord_cartesian(xlim = NULL, ylim = (c(0, 1))) +
  labs(title = "Largest patch index")

ggplot(classStats %>% filter(name == "core area percentage of landscape")) +
  geom_col(aes(x = flNm, y = value, fill = US_L3NAME)) +
  facet_wrap(~name, scales = "free")



#number of patches / area



ggplot(patchStats, aes(x = US_L3NAME, y = value, fill = US_L3NAME)) +
  geom_boxplot() +
  facet_wrap(~name, scales = "free") +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="black", fill="black") +
  labs(title = "PatchStats All")

ggplot(classStats, aes(x = US_L3NAME, y = value, fill = US_L3NAME)) +
  geom_boxplot() +
  facet_wrap(~name, scales = "free") +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="black", fill="black") +
  labs(title = "ClassStats All")


ggplot(patchStats %>% filter(name == "patch area" | name == "core area"), aes(x = US_L3NAME, y = value)) +
  geom_boxplot(aes(color = US_L3NAME)) +
  facet_wrap(~name) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="black", fill="black") +
  coord_cartesian(xlim = NULL, ylim = (c(0, 25))) +
  labs(title = "Patch Areas") +
  theme(axis.text.x=element_blank())
ggsave(here::here('figs', 'patchAreasL3Lim.png'))


ggplot(patchStats %>% filter(name == "patch area" | name == "core area"), aes(x = US_L3NAME, y = value)) +
  geom_boxplot(aes(color = US_L3NAME)) +
  facet_wrap(~name) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="black", fill="black") +
#  coord_cartesian(xlim = NULL, ylim = (c(0, 2000))) +
  labs(title = "Patch Areas") +
  theme(axis.text.x=element_blank())
ggsave(here::here('figs', 'patchAreasL3NoLim.png'))

ggplot(patchStats %>% filter(name == "shape index"), aes(x = US_L3NAME, y = value)) +
  geom_boxplot(aes(color = US_L3NAME)) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="black", fill="black") +
  coord_cartesian(xlim = NULL, ylim = (c(0, 20))) +
  labs(title = "Shape index")




ggplot(patchStats %>% filter(name == "core area index"), aes(x = US_L3NAME, y = value)) +
  geom_boxplot(aes(color = US_L3NAME)) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="black", fill="black") +
  coord_cartesian(xlim = NULL, ylim = (c(0, 100))) +
  labs(title = "Core area index")
ggsave(here::here('figs', 'caIndexL3.png'))


#Function to filter a dataframe to only include values >= the selected percentile in each group
#Group is the column name to group by
#Value is the column with the values to calculate percentiles of
#Percentile is the percentile in the form e.g. 0.75
filter_by_percentile <- function(data, group, value, percentile) {
  result <- data %>%
    group_by({{group}}) %>%
    do(filter(., {{value}} >= quantile({{value}}, percentile))) %>%
    ungroup()
  return(result)
}


#Examine core area extremes

percentiles <- c(.8, .85, .9, .95, .975, .99)

patchAreaExtremes <- data.frame()

for(p in percentiles) {
  patchAreaFilt <- filter_by_percentile(data = patchStats %>% filter(name == "core area"),
                                    group = US_L3NAME,
                                    value = value,
                                    percentile = p) %>%
    mutate(Percent = p * 100)
  patchAreaExtremes <- rbind(patchAreaExtremes, patchAreaFilt)
}

ggplot(patchAreaExtremes, aes(x = US_L3NAME,
                          y = value)) +
  geom_boxplot(aes(color = US_L3NAME)) +
  stat_summary(fun = "mean", geom="point", shape=20, size=3, color="black", fill="black") +
  facet_wrap(~Percent) +
  labs(title = "Patch Core Area, top percentiles (no Y limit)") +
  xlab("EPA Lvl 3 Ecoregion") +
  theme(axis.text.x=element_blank())

ggplot(patchAreaExtremes, aes(x = US_L3NAME,
                          y = value)) +
  geom_boxplot(aes(color = US_L3NAME)) +
  stat_summary(fun = "mean", geom="point", shape=20, size=3, color="black", fill="black") +
  facet_wrap(~Percent) +
  coord_cartesian(xlim = NULL, ylim = (c(0, 1500))) +
  labs(title = "Patch Core Area, top percentiles (Y: 0-1500)") +
  xlab("EPA Lvl 3 Ecoregion") +
  theme(axis.text.x=element_blank())

ggsave(here::here('figs', 'patchAreaExtremes.png'))





# show_correlation(patchStats %>% filter(region == "Eastern Batholith") %>% select(-region), method = "pearson")
# 
# 
# show_correlation(patchStats %>% filter(region == "Lochsa-Selway-Clearwater Canyons") %>% select(-region), method = "pearson")


```

```{r}

classStatsReshape <- classStats %>%
  select(-type, -function_name, -metric) %>%
  pivot_wider(names_from = name, values_from = value)

patchStatsReshape <- patchStats %>%
  select(-type, -function_name, -metric) %>%
  pivot_wider(names_from = name, values_from = value)

patchSummary <- patchStatsReshape %>%
  group_by(flNm) %>%
  summarise(medCorePSize = median(`core area`),
            meanCorePSize = mean(`core area`),
            maxCorePSize = max(`core area`))



classStatsMap <- aoiL4Interest %>%
  left_join(nameJoin) %>%
  left_join(classStatsReshape,
            join_by(flNm == flNm,
                    US_L3NAME == US_L3NAME),
            keep = FALSE) %>%
  st_as_sf() %>%
  filter(flNm != "AlpineZone") %>%
  mutate(area = st_area(.)) %>%
  left_join(patchSummary) %>%
  left_join(forestL4Stats, by = join_by(flNm == flNm)) %>%
  mutate(nPatchPerFArea = `number of patches` / forestAreaHa,
        `core area percentage of forest` = (`total core area` / forestAreaHa) * 100)


tm <- tm_shape(classStatsMap) +
  tm_fill(col = "core area percentage of forest",
          colorNULL = "eeeeee",
          style = "cont") +
  tm_layout(title = "Core area percentage of the forest landscape")
tm
tmap_save(tm, here::here('figs', 'caPerL4Map.png'))

# tm <- tm_shape(classStatsMap %>% filter(!is.na(`core area percentage of forest`))) +
#   tm_fill(col = "core area percentage of forest",
#           colorNULL = "eeeeee",
#           style = "log10") +
#   tm_layout(title = "Core area percentage of the forest landscape")
# tm
# tmap_save(tm, here::here('figs', 'caPerL4MapLog10.png'))

tm <- tm_shape(classStatsMap %>% filter(`core area percentage of forest` < 2)) +
  tm_fill(col = "core area percentage of forest",
          colorNULL = "eeeeee",
          style = "cont") +
  tm_layout(title = "Core area percentage of the forest landscape <2%")
tm
tmap_save(tm, here::here('figs', 'caPerL4Map_noOutlier.png'))

tm <- tm_shape(classStatsMap) +
  tm_fill(col = "number of patches", colorNULL = "eeeeee") +
  tm_layout(title = "number of patches")
tm
tmap_save(tm, here::here('figs', 'npatchL4Map.png'))


tm <- tm_shape(classStatsMap) +
  tm_fill(col = "nPatchPerFArea", colorNULL = "eeeeee") +
  tm_layout(title = "Number of patches per forest area")
tm
tmap_save(tm, here::here('figs', 'npatchpAreaL4Map.png'))

tm_shape(classStatsMap) +
  tm_fill(col = "largest patch index", colorNULL = "eeeeee") +
  tm_layout(title = "largest patch index")

tm <- tm_shape(classStatsMap) +
  tm_fill(col = "medCorePSize", colorNULL = "eeeeee") +
  tm_layout(title = "Median core size")
tm
tmap_save(tm, here::here('figs', 'medCorePL4Map.png'))

tm <- tm_shape(classStatsMap) +
  tm_fill(col = "meanCorePSize", colorNULL = "eeeeee") +
  tm_layout(title = "Mean core size")
tm
tmap_save(tm, here::here('figs', 'meanCorePL4Map.png'))

tm <- tm_shape(classStatsMap) +
  tm_fill(col = "maxCorePSize", colorNULL = "eeeeee") +
  tm_layout(title = "Max core size")
tm
tmap_save(tm, here::here('figs', 'maxCorePL4Map.png'))


```

```{r}

ggplot(classStatsMap, aes(x = `core area percentage of forest`, y = meanCorePSize)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq() +
  labs(title = "Mean core size")

ggplot(classStatsMap, aes(x = `core area percentage of forest`, y = medCorePSize)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq() +
  labs(title = "Median core size")
ggsave(here::here('figs', 'MedianCoreSizeCAPF.png'))

ggplot(classStatsMap, aes(x = `core area percentage of forest`, y = maxCorePSize)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq() +
  labs(title = "Max core size")

ggplot(classStatsMap, aes(x = `core area percentage of forest`, y = nPatchPerFArea)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq() +
  labs(title = "Number of patches per forest area")
  


patchStatsL4Top90Carea <- filter_by_percentile(data = patchStatsReshape,
                                          group = flNm,
                                          value = `core area`,
                                          percentile = .9) %>%
  group_by(flNm) %>%
  summarize(medCorePSize90 = median(`core area`),
            meanCorePSize90 = mean(`core area`))

patchStatsL4Top95Carea <- filter_by_percentile(data = patchStatsReshape,
                                          group = flNm,
                                          value = `core area`,
                                          percentile = .95) %>%
  group_by(flNm) %>%
  summarize(medCorePSize95 = median(`core area`),
            meanCorePSize95 = mean(`core area`))

patchStatsL4Top975Carea <- filter_by_percentile(data = patchStatsReshape,
                                          group = flNm,
                                          value = `core area`,
                                          percentile = .975) %>%
  group_by(flNm) %>%
  summarize(medCorePSize975 = median(`core area`),
            meanCorePSize975 = mean(`core area`))

classStatsMap2 <- classStatsMap %>% 
  left_join(patchStatsL4Top90Carea,
           join_by(flNm == flNm)) %>% 
    left_join(patchStatsL4Top95Carea,
           join_by(flNm == flNm)) %>%
  left_join(patchStatsL4Top975Carea,
           join_by(flNm == flNm))

ggplot(classStatsMap2, aes(x = `core area percentage of forest`, y = medCorePSize90)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq() +
  labs(title = "Median core size, 90 percentile")
ggsave(here::here('figs', 'MedianCoreSizeCAPF90.png'))

ggplot(classStatsMap2, aes(x = `core area percentage of forest`, y = meanCorePSize90)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq() +
  labs(title = "Mean core size, 90 percentile")

ggplot(classStatsMap2, aes(x = `core area percentage of forest`, y = medCorePSize95)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq() +
  labs(title = "Median core size, 95 percentile")
ggsave(here::here('figs', 'MedianCoreSizeCAPF95.png'))

ggplot(classStatsMap2, aes(x = `core area percentage of forest`, y = meanCorePSize95)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq() +
  labs(title = "Mean core size, 95 percentile")

ggplot(classStatsMap2, aes(x = `core area percentage of forest`, y = medCorePSize975)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq() +
  labs(title = "Median core size, 97.5 percentile")
ggsave(here::here('figs', 'MedianCoreSizeCAPF975.png'))

ggplot(classStatsMap2, aes(x = `core area percentage of forest`, y = meanCorePSize975)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq() +
  labs(title = "Mean core size, 97.5 percentile")

ggplot(classStatsMap2, aes(x = `core area percentage of forest`, y = `largest patch index`)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq() +
  labs(title = "Largest Patch Index")


```

```{r}
tm_shape(classStatsMap2) +
  tm_fill(col = "meanCorePSize90", colorNULL = "eeeeee") +
  tm_layout(title = "Mean core size, 90th percentile")

l3cap <- patchStatsReshape %>%
  group_by(US_L3NAME) %>%
  summarize(totCA = sum(`core area`)) %>%
  left_join(aoiL3Interest %>% 
    mutate(tAreaHa = drop_units(st_area(aoiL3Interest)) * .0001) %>%
    mutate(US_L3NAME = NA_L3NAME)) %>%
  mutate(`core area percentage of landscape` = (totCA / tAreaHa) * 100) %>%
  left_join(forestL3Stats) %>%
  mutate(`core area percentage of forest` = (totCA / forestAreaHa) * 100) %>%
    mutate(US_L3NAME = factor(US_L3NAME, level = c('Middle Rockies',
                                                 'Southern Rockies',
                                                 'Idaho Batholith',
                                                 'Sierra Nevada')))

ggplot(l3cap) +
  geom_col(aes(x = US_L3NAME, y=`core area percentage of forest`, fill = US_L3NAME)) +
  labs(title = "Fire/Insect Core Area percentage of Forested (BPS) Landscape")
ggsave(here::here('figs', 'caPerL3.png'))


